Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-06-17 SIR_manuscript.tex
Comments and edits focused on middle of paper

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}

@@

The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics. Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent), testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates. We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

@@

Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
primarily diagnostic (determining infection status for clinical purposes), or for control (determining  infection status in order to isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).

@@

We developed a deterministic model, Eqs.~\eqref{model}, which groups individuals based on disease status and testing status. Disease states include Susceptible, Infectious and Recovered; testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}). The testing status of an individual in  a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
Further, two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests. The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (Sec. \ref{app:R0}).

@@

In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W},
\end{equation}
where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
Thus, the \percap testing rate for compartment $X$ is 
\begin{equation}
\label{F}
F_X=\sigma w_X, \quad \text{where} \quad X \in \{S,I,R\}.
\end{equation}
For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.  Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate. Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (in the appendix we present an adjustment to the model that solves this problem).

The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}. In addition to these standard assumptions, our model assumes: 
\begin{enumerate}[(i)]
\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as
  \begin{equation}
  \label{Lambda}
  \Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
  \end{equation}
  with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a positive test (``confirmed positive'': Table~\ref{tab:params}). Susceptible individuals who are waiting for test results experience an additional transmission factor of $1-\theta\_w$ (\fref{fig:flowchart}). 
\item confirmed-positive individuals isolate more effectively than those awaiting test results, i.e.
$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
\end{enumerate}
For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$). Thus there are no waiting-for-positive or confirmed-positive susceptible individuals (we assume that these states are also absent from the initial conditions, i.e. $S\_p(0)=S\_c(0)=0$), which reduces the number of model states from 12 to 10.

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $w_S, w_I, w_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model \eqref{model}.}
\end{table}

@@

The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ as
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
\end{equation}
where the term $\beta/\gamma$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and $\Delta>0$ is the effectiveness of control due to testing and isolation processes defined as 
\begin{equation}
  \label{eq:del4a}
  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
\end{equation}
where
\begin{align}
\label{eq:C12}
C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
\end{align}
(Appendix~\ref{app:R0} gives a detailed derivation of these expressions).
The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$. In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (App.~\ref{app:rho}). 
We derived general expressions for these derivatives. However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity. Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.

The computations in this paper, specifically, calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}. We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{eq:del4a}) over a range of parameters.
We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).

The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size. Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$. This testing intensity reflects realistic testing rates during the \covid pandemic, i.e. a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021). The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic. The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.

@@

We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic, i.e. when the partial derivative of $\Delta$ has a consistent sign
(Appendices~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).

@@

Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions. In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g. symptomatic people or close contacts of known infections).

However, we did find two surprising trends: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.

@@

While slowing test returns does decrease $\Rnum$ over a broad range of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure. First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e. detecting and containing outbreaks while they are still in progress. This process could be modeled phenomenologically by making the testing focus an increasing function of the speed of test returns (i.e. $\partial w_I(\omega)/\partial \omega > 0$), because rapid testing allows more tests to be concentrated on infected individuals. Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns an increasing function of test return speed ($\partial \theta\_w(\omega)/\partial \omega > 0$), or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate. Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo: masking, ventilation, distancing measures, retail and event closures, stay-at-home orders, \ldots

@@

It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
\begin{align}
\label{cond}
& \omega>0, \\
& 0 \leq \rho<\omega,\\ 
& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
& \frac{w_I}{w_S}\geq 1.
\end{align}
In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}). 
In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.

@@

% %%%%%%%
\subsection{On Testing Rate and Numerical Singularity}

>>>>

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\appref}[1]{Sec.~\ref{app:#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}

@@

The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting). In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics. Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates. We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

@@

Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
primarily diagnostic (determining infection status for clinical purposes), or for control (determining  infection status in order to isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
In these cases, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).

@@

We developed a deterministic model which groups individuals based on disease status and testing status. Disease states include Susceptible, Infectious and Recovered; testing states include \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}). The testing status of an individual in  a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests. The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in Sec. \appref{R0}.

@@

In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
\begin{equation}
\label{sigma}
\sigma = \frac{\rho N_0}{W},
\end{equation}
where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
\begin{equation}
\label{F}
F_X=\sigma w_X.
\end{equation}
For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.  Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate. Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in Sec \appref{singularity}).

The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}. In addition to these standard assumptions, our model assumes: 
\begin{enumerate}[(i)]
\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as
  \begin{equation}
  \label{Lambda}
  \Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
  \end{equation}
  with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}). Susceptible individuals who are waiting for test results experience an additional transmission factor of $1-\theta\_w$ (\fref{fig:flowchart}). 
\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.
$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
\end{enumerate}
For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$). Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.

\begin{table}[htp]
\centering
\begin{tabular}{|c|L{2in}cc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
  $w_S, w_I, w_R$ & Relative testing weights & - &
  \begin{minipage}[t]{0.21\columnwidth}%
    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
  \end{minipage} \\
  \hline
  \end{tabular}
\caption{\label{tab:params} Parameters of the model specified in \eqref{model}; see also \fref{fig:flowchart}.}
\end{table}

@@

The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ as
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
\end{equation}
where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction due to testing and isolation processes, calculated as 
\begin{equation}
  \label{eq:del4a}
  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
\end{equation}
where
\begin{align}
\label{eq:C12}
C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
\end{align}
(\appref{R0} gives a detailed derivation of these expressions).
This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$. In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (\appref{rho}). 
We derived general expressions for these derivatives. However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity. Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.

Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).

\jd{Why do we use a different range of $\omega$ for the two cases?}
The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size. Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$. This testing intensity reflects realistic testing rates during the \covid pandemic, i.e. a maximum of 1.3\% of the population per day, giving a maximum of 13000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021). The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic. The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3.

@@

\jd{Should we talk less about \Rnum\ and keep the focus on $\Delta$? The back and forth is a bit tiring.}

We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic, i.e. when the partial derivative of $\Delta$ has a consistent sign
(see Appendices).

@@

Many of the qualitative results we have derived confirm simple, common-sense intuitions. In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g. symptomatic people or close contacts of known infections).

However, we did find two surprising patterns: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.

@@

While slowing test returns does decrease $\Rnum$ over a broad range of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure. First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e. detecting and containing outbreaks while they are still in progress. This process could be modeled phenomenologically by making the testing focus increase as an increasing proportion of cases are detected, because finding infections allows tests to be concentrated on their connections. Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate. Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo -- these include masking, ventilation, distancing measures, retail and event closures, or stay-at-home orders.

@@

It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
\begin{align}
\label{cond}
& \omega>0, \\
& 0 \leq \rho<\omega,\\ 
& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
& \frac{w_I}{w_S}\geq 1.
\end{align}
In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}). 
In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.

@@

% %%%%%%%
\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
